TimeVaryingDetection_mrmr

### 2024-11-21 code/R/54188_ClassicCustomComparison_mrmr.Rmd
continue writing model block sections

The R package mrmr serves as an interface to prepare data and run population models. It relies on a site-specific multi-state Bayesian open-population Jolly-Seber CMR model (hereafter referred to as the multi-state model) for the mathematical and computational part of model fitting.The multi-state model is written in the Stan programming language, which is designed for probabilistic modeling and Bayesian inference. The 'changes to mrmr' outlined in this report are primarily changes to the multi-state model, via modifications to the Stan code. The Stan code is not included in this report. Instead, changes are described using the English language. Snippets of essential Stan code are provided, when necessary, in *italics*.

@@

4.  Parameters - declared the parameters the model will estimate.

5.  Transformed parameters

@@

7.  Generated quantities - calculates quantities after the model is fit, used for interpretation and diagnostics.

Changes made to each block are described below.

@@

The changes made to this block are essential to achieving independent, time-varying estimates of detection probability.

**beta_detect** a vector of parameters that represent the logit-scale detection probability for survey sessions. The classic mrmr model defines beta_detect as a vector of size m_detect. m_detect refers to the number of detection covariates. The custom model defines beta_detect as a vector of size Jtot. Jtot refers to the total number of survey sessions.

Simply, the original line *vector[m_detect] beta_detect*; is changed to *vector[Jtot] beta_detect;* where each entry corresponds to a specific survey session.

Meaning is assigned to beta_detect in the functions block via transformed parameter logit_detect, within the forward_prob( ) function. No changes were made to this function.

@@

In the classic model, logit_detect is derived from beta_detect using the design matrix x_detect. This ties detection probabilities to shared covariates. In the custom model, logit_detect is assigned directly to beta_detect, making each session's detection probability independent. \*

-   When I want to introduce covariates into my analysis, I will need to revisit this section.

### The data

@@

## Run Models

@@

## Model Comparison

### Generated Quantities

@@

### Quantifed Model Fit

>>>>

The R package mrmr serves as an interface to prepare data and run population models. It relies on a site-specific multi-state Bayesian open-population Jolly-Seber CMR model (hereafter referred to as the multi-state model) for the mathematical and computational part of model fitting.The multi-state model is written in the Stan programming language, which is designed for probabilistic modeling and Bayesian inference. The 'changes to mrmr' outlined in this report are primarily changes to the multi-state model, via modifications to the Stan code. The Stan code is not included in this report. Instead, changes are described using the English language. Snippets of essential Stan code are provided, when necessary, in *italics*, following the format (*classic model code*) —\> (*custom model code*).

@@

4.  **Parameters** - declared the parameters the model will estimate.

5.  **Transformed parameters**

@@

7.  **Generated quantities** - calculates quantities after the model is fit, used for interpretation and diagnostics.

This report focuses on changes made to three of those blocks: generates quantities, parameters, and transformed parameters. Those changes, and their functionality, are outlined below.

@@

(*vector[m_detect] beta_detect;*) —\> (*vector[Jtot] beta_detect;*)

This change is essential to achieving independent, time-varying estimates of detection probability in the custom model.

**beta_detect** is a vector of parameters that represent the logit-scale detection probability for survey sessions. The classic mrmr model defines beta_detect as a vector of size m_detect. m_detect refers to the number of detection covariates. The custom model defines beta_detect as a vector of size Jtot. Jtot refers to the total number of survey sessions. Each entry in the vector now corresponds to a specific survey session (custom) rather than a covariate (classic).

Meaning is assigned to beta_detect in the functions block via the transformed parameter logit_detect, within the forward_prob( ) function. No changes were made to this function.

@@

In the classic model, logit_detect is derived from beta_detect using the design matrix of covariates x_detect. This ties detection probabilities to shared covariates. In the custom model, logit_detect is assigned directly to beta_detect, making each session's detection probability independent.

Changes to the parameters and transformed parameters blocks reflect a purpose to obtain the most accurate estimates of N, B, p, and phi for the collective population. The classic model seems to reflect a purpose to model detection and survival by covariates, which makes sense in the context of its original purpose, described by Knapp and Joseph (2018). If and when I want to introduce covariates into my analysis, I will need to revisit this section. (I'm not sure if it is possible to include both covariates in the design matrix and indepentent, session-specific detection.)

## The Data

@@

## Analysis

This analysis serves to...

@@

### Model Comparison

#### Generated Quantities

@@

#### Model Fit
